---
title: "Tooth Growth Analysis"
author: "Victor Rainier Cruz Perez"
date: "September 20th, 2014"
output: pdf_document
---


### Exploring Tooth Growth Data
```{r}
library(datasets);data(ToothGrowth);library(ggplot2)
data <- ToothGrowth

#loading required libraries
require(grid)
#defining layout
vp.layout <- function(x, y) viewport(layout.pos.row=x, layout.pos.col=y)
arrange_ggplot2 <- function(..., nrow=NULL, ncol=NULL, as.table=FALSE) {
           dots <- list(...)
              n <- length(dots)
          	
            if(is.null(nrow) & is.null(ncol)) { nrow = floor(n/2) ; ncol = ceiling(n/nrow)}
          	if(is.null(nrow)) { nrow = ceiling(n/ncol)}
          	if(is.null(ncol)) { ncol = ceiling(n/nrow)}

           
grid.newpage()
pushViewport(viewport(layout=grid.layout(nrow+1,ncol,heights = unit(c(0.5,5),"null")  ) ))
	  ii.p <- 1
	  for(ii.row in seq(1, nrow)){
	    ii.table.row <- ii.row	
	    if(as.table) {ii.table.row <- nrow - ii.table.row + 1}
		    for(ii.col in seq(1, ncol)){
			    ii.table <- ii.p
			    if(ii.p > n) break
			    print(dots[[ii.table]], vp=vp.layout(ii.table.row+1, ii.col))
			    ii.p <- ii.p + 1}}}
```

```{r}
## Ploting the dose factor
ch1<-ggplot(data,aes(x=factor(dose),y=len,fill=factor(dose)))+
      geom_boxplot(notch=T,notchwidth = 0.2)+theme_bw()

## Ploting the dose vs Teeth's length
ch1 <- ch1+facet_wrap(~supp)+ theme(legend.position="none")+
      scale_x_discrete("Dosage in mg")+scale_y_continuous("Length of Teeth")

## Ploting the the linear regression 
ch2<-ggplot(data,aes(x=dose,y=len,color=supp,group=supp))+
      geom_point()+stat_smooth(method="lm",alpha=0.1)


ch2 <-ch2+theme_bw()+theme(legend.justification=c(1,0),legend.position=c(1, 0))+
      scale_color_discrete("Supplement")

ch2<-ch2+scale_x_continuous("Dosage in mg")+scale_y_continuous("Length of Teeth")
```


###Summary of Data Analysis (Tooth Growth Data and Exploratory )
```{r}
library(dplyr);library(reshape2);library(knitr)
aux <- data %.% group_by(supp,dose) %.% summarise(Mean=mean(len),
                                                  Min=min(len),
                                                  Max=max(len),"Std. Dev"=sd(len))
aux2<-dcast(melt(aux,id.vars=c("supp","dose")),dose~supp+variable,fun.aggregate=sum)
aux2<- round(aux2,2)
names(aux2)[1] <- "Dosage"
kable(aux2, format = "markdown")
```

The table above shows the sumamry statistics for the Tooth Growth data. One can infer, that orange juice is more effective at lower dosages; but that BTH supplements are similalry effective at the 2mg dose. Also one can infer that Vitamon C has less variability forlower doses and more than at higher dose.



```{r,warning=F}
arrange_ggplot2(ch1,ch2)
grid.text("Exploratory Analysis for Tooth Growth Data", 
          vp = viewport(layout.pos.row = 1, 
                        layout.pos.col = 1:2))
```

The figure above shows the impact of the supplement and the dosage on the length of teeth of Guinea Pigs. 

The left panel shows the distributions for different combinations of supplement and dosage. From the analysis of the left panel, one can assume that their medians are different.

The right panel shows a scatter plot with linear regression where the teeth length fits as a function of dosage for each supplement.

From the analysis of both Panels, one can infer that higher dosages appears to be related to longer teeth. 


###Compare Peformance and Dosage
* Is Orange Juice more effective than Vitamin C?

```{r}
t.test(len~supp,paired=F,var.equal=T,data=data,alternative="g")$p.value
t.test(len~supp,paired=F,var.equal=F,data=data,alternative="g")$p.value

#adjusted values
t.test(len~supp,paired=F,var.equal=T,data=data,alternative="g")$p.value*50
t.test(len~supp,paired=F,var.equal=F,data=data,alternative="g")$p.value*50
```
After performing a t-test in overall orange juice had a greater impact on tooth growth than Vitamin C.  However the use of Bonferroni adjusted values suggest that is not possible to affirm that there is no difference in the 2 supplements

* Does Dosage Matter?

```{r}
results = data.frame(supplement=character(),dose_1=numeric(),
                     dose_2=numeric(),ll=numeric(),
                     ul=numeric(),pval=numeric(),
                     variance_equal=logical(),
                     test=character(),stringsAsFactors=F)
k=1
for (i in levels(data$supp)) {
  for (j in unique(data$dose)) {
    for (test in c("greater","less")) {
      for (variance in c(T,F)) {
        data1 <- subset(data,supp==i & dose!=j)
        temp <- t.test(len~dose,paired=F,var.equal=variance,data=data1,alternative=test)
        results[k,"supplement"] <- i; results[k,"dose_1"] <- unique(data1$dose)[1]; 
        results[k,"dose_2"] <- unique(data1$dose)[2]; results[k,"ll"] <- temp$conf.int[1]
        results[k,"ul"] <- temp$conf.int[2]; results[k,"pval"] <- temp$p.value
        results[k,"variance_equal"] <- variance; results[k,"test"] <- test
        k <- k+1
      }}}}

results$adj_pval <- results$pval*50
results$outcome <- ifelse(results$pval<=0.05,"Difference","No Difference")
results[c(4:6,9)] <- round(results[c(4:6,9)],3)
results2 <- subset(results,outcome=="Difference")
results2 <- results2[,c(1:6,9,7,8,10)]
kable(results2, format = "markdown")
```

The table above shows the dosage comparison for each supplement. It shows that higher dosages are more effective that lower dosages for either supplement. However, after the use the Bonferroni adjustment for p-values one cannot conclude that the 2mg dosage for orange juice is more effective than the 1mg dosage.



* Which Supplement is Better at a Given Dosage?
```{r,echo=TRUE,results='asis',warning=F,message=F,tidy=F,fig.align='center'}
results_a = data.frame(dose=numeric(),supplement1=character(),supplement2=character(),ll=numeric(),
                     ul=numeric(),pval=numeric(),variance_equal=logical(),test=character(),stringsAsFactors=F)
k=1
for (j in unique(data$dose)) {
    for (test in c("greater","less")) {
      for (variance in c(T,F)) {
        data1 <- subset(data, dose==j)
        temp <- t.test(len~supp,paired=F,var.equal=variance,data=data1,alternative=test)
        results_a[k,"supplement1"] <- levels(data1$supp)[1]; results_a[k,"supplement2"] <- levels(data1$supp)[2]
        results_a[k,"dose"] <- j; results_a[k,"ll"] <- temp$conf.int[1]
        results_a[k,"ul"] <- temp$conf.int[2]; results_a[k,"pval"] <- temp$p.value
        results_a[k,"variance_equal"] <- variance; results_a[k,"test"] <- test
        k <- k+1
      }
    }
}

results_a$adj_pval <- results_a$pval*50
results_a$outcome <- ifelse(results_a$pval<=0.05,"Difference","No Difference")
results_a[c(4:6,9)] <- round(results_a[c(4:6,9)],3)
results_a2 <- subset(results_a,outcome=="Difference")
results_a2 <- results_a2[,c(1:6,9,7:8,10)]
kable(results_a2, format = "markdown")
```

The table above shows the supplement comparisons for each dosage for which a statistical difference at 95% was observed. It shows that orange juice is more effective at 0.5 and 1 mg but that both supplements are equally effecive at 2mg. When we use the Bonferroni adjustment on the p-values, we can only conclude that orange juice is more effective at 1mg

* Which Dosage is Better?
```{r}
results_b = data.frame(dose1=numeric(),dose2=numeric(),
                       ll=numeric(),ul=numeric(),pval=numeric(),
                       variance_equal=logical(),test=character(),
                       stringsAsFactors=F)
k=1
for (j in unique(data$dose)) {
    for (test in c("greater","less")) {
      for (variance in c(T,F)) {
        data1 <- subset(data, dose!=j)
        temp <- t.test(len~dose,paired=F,var.equal=variance,data=data1,alternative=test)
        results_b[k,"dose1"] <- unique(data1$dose)[1]; results_b[k,"dose2"] <- unique(data1$dose)[2]
        results_b[k,"ll"] <- temp$conf.int[1]; results_b[k,"ul"] <- temp$conf.int[2]; 
        results_b[k,"pval"] <- temp$p.value
        results_b[k,"variance_equal"] <- variance; results_b[k,"test"] <- test
        k <- k+1
      }}}

results_b$adj_pval <- results_b$pval*50
results_b$outcome <- ifelse(results_b$pval<=0.05,"Difference","No Difference")
results_b[c(3:5,8)] <- round(results_b[c(3:5,8)],3)
results_b$difference <- ifelse(results_b$outcome=="Difference",T,F)
results_b2 <- subset(results_b,outcome=="Difference")
results_b2 <- results_b2[c(1:5,8,6,7,9:10)]
kable(results_b2, format = "markdown")

```

The table above shows the comparison across supplements for each dosage.It shows that 2mg is superior to 1 and 0.5 mgs, and that 1mg is superior to 0.5 mg. After the use of Bonferroni adjustmnt on the p-values, the results where confirmed.


